Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2014-12-08 fib2.tex
some edits and a ref added.

%------------------------------------------------------------------------------------------------------------%
\paragraph{Modelling Fibonacci anyons}

A physical anyon model is abstractly described by a unitary modular tensor 
category~\cite{Wang2010b}. This object contains the data describing the types of anyonic 
particles, as well as the results of topological operations such as braiding or fusion of these 
anyons. The Fibonacci anyon model consists of only one non-trivial particle type, 
conventionally called $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the 
possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two 
$\tau$ particles may either fuse to vacuum or to a $\tau$ particle.

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the anyon 
model correspond to large length-scale physics, while the microscopic physics plays a less 
important and non-universal role. As such, we model the system simply as a square lattice of 
sites on which sets of anyons may reside. Allowed anyon dynamics are considered to be local 
on this lattice, which we take to have periodic boundary conditions. For more details of an 
analogous model, see~\cite{Brell2013}.

%------------------------------------------------------------------------------------------------------------%
\paragraph{Noise and error-correction}

@@

\cggb{Details. Consider merging parts of this section with parts of \Sref{s:simalg}.}

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Logical errors}

@@

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Decoding algorithm}

@@

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Simulation}

Although simulation of general processes involving $n$ Fibonacci anyons is BQP-complete (and thus unlikely to be classically tractable), noise and error-correction procedures have structure that we can exploit to efficiently simulate typical processes of interest. In particular, those processes in which we expect error-correction to succeed are also those that we expect to be able to efficiently simulate for the following reasons.

Below the (bond) percolation threshold, we expect random sets of bonds to decompose into separate clusters of average size $O(log(n))$ and variance $O(1)$, \cite{Bazant2000}. In our context, the random bonds correspond to noise processes, and so separate clusters correspond to sets of anyons that could not have interacted at any point in their history. As such, we can simulate the braiding processes of each cluster separately. In other words: it is a (tensor) product state, so we compute with each factor separately. Since each cluster has only size $O(log(n))$, we may typically simulate these dynamics efficiently.

However, random noise processes are not the only dynamics that we need consider. We must also consider the effect of the error-correction routine itself. This acts iteratively to fuse particles on increasing length scales. While this kind of fusion would typically merge clusters, forcing us to compute dynamics of larger and larger sets of anyons, at each length scale the total number of anyons present is dramatically reduced. Indeed, the goal of the error correction algorithm is to reduce the number of anyons as much as possible.

\cggb{Can we make a concrete claim about the maximum size of clusters at each stage of the decoding procedure? Do we know anything about the distribution of separations between clusters in the sub-critical percolation regime?}

Of course, if we were to consider noise so strong that the clusters percolate over the lattice, we would no longer be able to efficiently simulate the system. However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. If the noise operations are so common to have percolated, this is an example of a situation where we expect error-correction \emph{not} to succeed. Similarly, if the noise has not yet percolated by itself, but the clusters are merged into a large (percolated) cluster by the decoding algorithm, this is again an example of a process that we expect to correspond to a failure of error-correction. So we see that the events that we are able to efficiently simulate correspond precisely to those for which we expect error-correction to succeed. We may simply declare failure on non-simulable events. This allows us to simulate and demonstrate the success of error-correction schemes for Fibonacci anyons, without explicitly simulating the failure of such schemes.

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Simulation algorithm}

We cover the surface of the torus with LxL disjoint squares. These {\it tiles} are the regions over which
charge measurements occur, in order to produce a syndrome for the decoder.

In order to keep track of the observables of the system we maintain a set of disjoint {\it curves}. These curves can also be thought of as a top-down view of the fusion tree. In general, this winds around the two dimensional tiles in a haphazard way determined by the progress of the simulation. For comparison, previous work used a single fixed Z-shaped curve (row-major ordering) \cite{Brell2013}.

\begin{center}
	{\scriptsize (i)
	\includegraphics[width=0.26\columnwidth]{pair-create.pdf}
	\hskip 5pt
	(ii)
	\includegraphics[width=0.26\columnwidth]{syndrome-1.pdf}
	\hskip 5pt
	(iii)}
	\includegraphics[width=0.26\columnwidth]{syndrome-2.pdf}
\end{center}

@@

%------------------------------------------------------------------------------------------------------------%
\paragraph{Numerical results}

@@

\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\end{center}

@@

%------------------------------------------------------------------------------------------------------------%
\paragraph{Discussion}

>>>>

%------------------------------------------------------------------------------------------------------------%
\paragraph{Modelling Fibonacci anyons.}

A physical anyon model is abstractly described by a unitary modular tensor 
category~\cite{Wang2010b}. This object contains the data describing the types of anyonic 
particles, as well as the results of topological operations such as braiding or fusion of these 
anyons. The Fibonacci anyon model consists of only one non-trivial particle type, 
conventionally called $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe the 
possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, i.e.~two 
$\tau$ particles can either fuse to vacuum or to a $\tau$ particle.

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the anyon 
model correspond to large length-scale physics, while the microscopic physics plays a less 
important and non-universal role. As such, we model the system simply as a square lattice of 
sites on which sets of anyons can reside. Allowed anyon dynamics are considered to be local 
on this lattice, which we take to have periodic boundary conditions. For more details of an 
analogous model, see~\cite{Brell2013}.

%------------------------------------------------------------------------------------------------------------%
\paragraph{Noise and error-correction.}

@@

\cggb{Details. Consider merging parts of this section with parts of Simulation below.}

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Logical errors.}

@@

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Decoding algorithm.}

@@

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Simulation.}

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
to in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise and error-correction procedures have structure that we can exploit to 
efficiently simulate typical processes of interest. In particular, those processes in which we 
expect error-correction to succeed are also those that we expect to be able to efficiently 
simulate for the following heuristic reasons.

Below the (bond) percolation threshold for (say) a square lattice, we expect random sets of 
bonds to decompose into separate clusters of average size $O(\log(n))$ and variance 
$O(1)$,~\cite{Bazant2000}. In our context, the random bonds correspond to noise processes, 
and so disconnected clusters correspond to sets of anyons that could not have interacted at 
any point in their history. As such, we can simulate the braiding processes of each cluster 
separately. In other words: the quantum state factorizes into a tensor product, and we can 
compute with each factor separately. Since each cluster has size only $O(\log(n))$, we can 
typically simulate these dynamics efficiently because the resulting fusion space has 
dimension $O(\mathrm{poly}(n))$. There are rare random processes that violate this 
reasoning, but processes such as these are suppressed exponentially in $L$, the 
lattice size~\cite{Grimmett1989}.

However, random noise processes are not the only dynamics that we need to consider. We 
must also consider the effect of the error-correction routine itself. This acts iteratively to fuse 
particles on increasing length scales. While this kind of fusion would typically merge clusters, 
forcing us to compute dynamics of larger and larger sets of anyons, at each length scale the 
total number of anyons present is dramatically reduced. Indeed, the goal of the error 
correction algorithm is to reduce the number of anyons as much as possible.

\cggb{Can we make a concrete claim about the maximum size of clusters at each stage of the decoding procedure?} \stf{Well, the worst case behavior has no bound, so we need a statement about expectation.} \cggb{Do we know anything about the distribution of separations between clusters in the sub-critical percolation regime?} \stf{I think the Bravyi \& Haah appendix might provide some results that would be useful in making such a statement.}

Of course, if we were to consider noise so strong that the clusters percolate over the lattice, we would no longer be able to efficiently simulate the system. However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. If the noise operations are so common to have percolated, this is an example of a situation where we expect error-correction \emph{not} to succeed. Similarly, if the noise has not yet percolated by itself, but the clusters are merged into a large (percolated) cluster by the decoding algorithm, this is again an example of a process that we expect to correspond to a failure of error-correction. So we see that the events that we are able to efficiently simulate correspond precisely to those for which we expect error-correction to succeed. We can simply declare failure on non-simulable events. This allows us to simulate and demonstrate the success of error-correction schemes for Fibonacci anyons, without explicitly simulating the failure of such schemes.

%------------------------------------------------------------------------------------------------------------%	
\paragraph{Simulation algorithm.}

We cover the surface of the torus with $L\times L$ disjoint squares. These \emph{tiles} are the regions over which
charge measurements occur, in order to produce a syndrome for the decoder.

In order to keep track of the observables of the system we maintain a set of disjoint \emph{curves}. These curves can also be thought of as a top-down view of the fusion tree. In general, this winds around the two dimensional tiles in a haphazard way determined by the progress of the simulation. For comparison, previous work used a single fixed Z-shaped curve (row-major ordering)~\cite{Brell2013}.

\begin{figure}[th!]
\begin{center}
	{\scriptsize (i)
	\includegraphics[width=0.27\columnwidth]{pair-create.pdf}
	\hskip 4pt
	(ii)
	\includegraphics[width=0.27\columnwidth]{syndrome-1.pdf}
	\hskip 4pt
	(iii)}
	\includegraphics[width=0.27\columnwidth]{syndrome-2.pdf}
\caption{default}
\label{default}
\end{center}
\end{figure}

@@

%------------------------------------------------------------------------------------------------------------%
\paragraph{Numerical results.}

@@

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\caption{default}
\label{default}
\end{center}
\end{figure}

@@

%------------------------------------------------------------------------------------------------------------%
\paragraph{Discussion.}
